Pollution, agricultural productivity, and development: Evidence from coal in plants in India\footnote{All scripts used in this paper are publicly available on my GitHub page: github.com/JoshMerfeld/pollution\_development}

### 2023-07-13 draft/draft.Rmd
introduction update

While there may be positive effects from this increase in power generation -- for example, electricity may lead to economic and productivity growth [@dinkelman2011effects; @kline2014local; @rud2012electricity; @van2017long], though electrification alone may not be a sufficient condition [@lee2020does] -- these improvements do not come without downsides, especially when it comes to power derived from coal. Pollutants from coal contribute to climate change and the phase out of coal energy is, therefore, a major climate goal [@iea2022]. Emissions from the production and burning of coal are also directly harmful to human health and the environment; pollutants like sulfur dioxide, nitrogen oxides, methane, and mercury are particularly harmful.[^4] While the use of coal had been declining, recent events -- including Russia's invasion of Ukraine -- have led to an increase in the demand for coal.[^5]

[^4]: www.eia.gov/energyexplained/coal/coal-and-the-environment.php

@@

@@

This paper contributes to several strands of literature. We already have evidence of effects of pollution on different forms of labor activity. For example, general levels of pollution decrease productivity of call center works in China [@chang2019effect], lead people to perform worse on cognitive functions [@ebenstein2016long; @wen2022lower], and also lead to lower levels of farm labor productivity in California, driven by changes in ozone concentration [@graff2012impact]. Also in China, @he2019severe find small but significant negative effects of pollution on manufacturing productivity, while @chen2022effect show pollution leads to labor out migration. Relevant to my results, @hanna2015effect show that decreases in pollution lead to increases in labor allocation in another developing country.

>>>>

While there may be positive effects from this increase in power generation -- for example, electricity may lead to economic and productivity growth [@dinkelman2011effects; @kline2014local; @rud2012electricity; @van2017long], though electrification alone may not be a sufficient condition [@lee2020does] -- these improvements do not come without downsides, especially when it comes to power derived from coal. Emissions from the production and burning of coal are also directly harmful to human health and the environment; pollutants like sulfur dioxide, nitrogen oxides, methane, and mercury are particularly harmful. While the use of coal had been declining, recent events -- including Russia's invasion of Ukraine -- have led to an increase in the demand for coal.[^5]

@@

Although we have made much progress over the last several decades, poverty and inequality remain important challenges in the developing world [@alvaredo2015recent]. Poverty is particularly prevalent in rural areas, where the majority of the world's poor live [@dercon2009rural], and in the agricultural sector, where the majority of the world's poor work [@christiaensen2011evolving]. As such, agricultural productivity is a key determinant of poverty in developing countries [@dercon2009rural], meaning that any threat to agricultural productivity is also a threat to poverty reduction.

@@

This paper contributes to several strands of literature. We already have evidence of effects of pollution on different forms of labor activity. For example, general levels of pollution decrease productivity of call center works in China [@chang2019effect], lead people to perform worse on cognitive functions [@ebenstein2016long; @wen2022lower], lower labor market earnings [@borgschulte2022air], and also lead to lower levels of farm labor productivity in California, driven by changes in ozone concentration [@graff2012impact]. Also in China, @he2019severe find small but significant negative effects of pollution on manufacturing productivity, while @chen2022effect show pollution leads to labor out migration. Relevant to my results, @hanna2015effect show that decreases in pollution lead to increases in labor allocation in another developing country.

### 2023-07-13 draft/draft.tex
introduction update

While there may be positive effects from this increase in power generation -- for example, electricity may lead to economic and productivity growth (Dinkelman 2011; Kline and Moretti 2014; Rud 2012; Van de Walle et al. 2017), though electrification alone may not be a sufficient condition (K. Lee, Miguel, and Wolfram 2020) -- these improvements do not come without downsides, especially when it comes to power derived from coal. Pollutants from coal contribute to climate change and the phase out of coal energy is, therefore, a major climate goal (IEA 2022). Emissions from the production and burning of coal are also directly harmful to human health and the environment; pollutants like sulfur dioxide, nitrogen oxides, methane, and mercury are particularly harmful.\footnote{www.eia.gov/energyexplained/coal/coal-and-the-environment.php} While the use of coal had been declining, recent events -- including Russia's invasion of Ukraine -- have led to an increase in the demand for coal.\footnote{www.npr.org/2022/08/15/1117560560/a-rising-demand-for-coal-amidst-war-in-ukraine}

@@

@@

This paper contributes to several strands of literature. We already have evidence of effects of pollution on different forms of labor activity. For example, general levels of pollution decrease productivity of call center works in China (Chang et al. 2019), lead people to perform worse on cognitive functions (Ebenstein, Lavy, and Roth 2016; Wen and Burke 2022), and also lead to lower levels of farm labor productivity in California, driven by changes in ozone concentration (Graff Zivin and Neidell 2012). Also in China, He, Liu, and Salvo (2019) find small but significant negative effects of pollution on manufacturing productivity, while Chen, Oliva, and Zhang (2022) show pollution leads to labor out migration. Relevant to my results, Hanna and Oliva (2015) show that decreases in pollution lead to increases in labor allocation in another developing country.

@@

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-iea2022}{}}%
IEA. 2022. {``{Coal-fired electricity}.''} {International Energy Association}.

>>>>

While there may be positive effects from this increase in power generation -- for example, electricity may lead to economic and productivity growth (Dinkelman 2011; Kline and Moretti 2014; Rud 2012; Van de Walle et al. 2017), though electrification alone may not be a sufficient condition (K. Lee, Miguel, and Wolfram 2020) -- these improvements do not come without downsides, especially when it comes to power derived from coal. Emissions from the production and burning of coal are also directly harmful to human health and the environment; pollutants like sulfur dioxide, nitrogen oxides, methane, and mercury are particularly harmful. While the use of coal had been declining, recent events -- including Russia's invasion of Ukraine -- have led to an increase in the demand for coal.\footnote{www.npr.org/2022/08/15/1117560560/a-rising-demand-for-coal-amidst-war-in-ukraine}

@@

Although we have made much progress over the last several decades, poverty and inequality remain important challenges in the developing world (Alvaredo and Gasparini 2015). Poverty is particularly prevalent in rural areas, where the majority of the world's poor live (Dercon 2009), and in the agricultural sector, where the majority of the world's poor work (Christiaensen, Demery, and Kuhl 2011). As such, agricultural productivity is a key determinant of poverty in developing countries (Dercon 2009), meaning that any threat to agricultural productivity is also a threat to poverty reduction.

@@

This paper contributes to several strands of literature. We already have evidence of effects of pollution on different forms of labor activity. For example, general levels of pollution decrease productivity of call center works in China (Chang et al. 2019), lead people to perform worse on cognitive functions (Ebenstein, Lavy, and Roth 2016; Wen and Burke 2022), lower labor market earnings (Borgschulte, Molitor, and Zou 2022), and also lead to lower levels of farm labor productivity in California, driven by changes in ozone concentration (Graff Zivin and Neidell 2012). Also in China, He, Liu, and Salvo (2019) find small but significant negative effects of pollution on manufacturing productivity, while Chen, Oliva, and Zhang (2022) show pollution leads to labor out migration. Relevant to my results, Hanna and Oliva (2015) show that decreases in pollution lead to increases in labor allocation in another developing country.

@@

\leavevmode\vadjust pre{\hypertarget{ref-alvaredo2015recent}{}}%
Alvaredo, Facundo, and Leonardo Gasparini. 2015. {``Recent Trends in Inequality and Poverty in Developing Countries.''} \emph{{Handbook of Income Distribution}} 2: 697--805.

@@

\leavevmode\vadjust pre{\hypertarget{ref-borgschulte2022air}{}}%
Borgschulte, Mark, David Molitor, and Eric Yongchen Zou. 2022. {``{Air pollution and the labor market: Evidence from wildfire smoke}.''} \emph{{Review of Economics and Statistics}}, 1--46.

@@
